### 2011-12-21 trunk/GUIErrorDetector/paper/abstract.tex
paper draft

\begin{abstract}
Graphical User Interface (GUI) is a critical component of today's software.
 It provides user-friendly access to the functionality provided by the application.
To keep GUIs responsive and active, a GUI application often spawns
 separate threads to handle lengthy operations in the background,
such as expensive computations, I/O tasks, and network requests.
However, many modern GUI frameworks require all GUI objects
must be accessed exclusively by the \textit{event dispatching thread}
(or called \textit{UI thread}). If a GUI object
is accessed through a non-UI thread, an \textit{invalid thread access} error
occurs and the whole application may abort consequently.

In this paper, we present a general technique to find such \textit{invalid thread access}
errors in multithreaded GUI applications. We formulate finding invalid
thread access errors as a call graph
reachability problem. Our technique starts from call graph entry nodes and statically explores
all possible paths that may cause a non-UI thread to access a GUI object.
However, the heavy use of reflection in modern GUI frameworks makes existing
call graph construction algorithms insufficent to build a both complete and precise
call graph. To overcome such limitations, we present a combined RTA and k-CFA
call graph analysis (that XXX). % to overcome such limitations.% of analyzing GUI applications.

We implemented our technique and instantiated it on four popular GUI
platforms, namely SWT, Eclipse plugin development framework, Java Swing, and Android. In an evaluation
on XX subjects over XXX LOC, our technique found a number of errors,
suggesting its effectiveness and showing good scalability.

\end{abstract}